---
title: "STAT545 Assignment 4: Tidy Data and Joins"
output:
  html_document:
    toc: yes
    keep_md: yes
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, echo = FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(tsibble))
```

# Part 1 - Univariate Option 1: Comparing Life Expectancies over Time

Full dataframe for reference...

```{r}
gapminder
```

## 1.1: Pivoting Wider to Comapre two Countries

```{r}
gapminder %>% 
  filter(country == "China" | country == "Taiwan") %>% #isolating counries of interest
  select(country, year, lifeExp) %>% #isolating variables of interest
  pivot_wider(id_cols = year, #pivot, keeping year column
              names_from = country, #pulling new columns from country column
              values_from = lifeExp) #pulling values from lifeExp column
```

## 1.2: Plotting with the New Data Layout

```{r}
gapminder %>% 
  filter(country == "China" | country == "Taiwan") %>% 
  select(country, year, lifeExp) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp) %>% 
  ggplot(aes(Taiwan, China)) + #plotting Taiwan against China
  geom_point(colour = "dark red") + #using point geom with dark red colour
  theme_minimal()
```

## 1.3: Pivoting Longer to Return Original Organization
 
```{r}
gapminder %>% 
  filter(country == "China" | country == "Taiwan") %>% 
  select(country, year, lifeExp) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp) %>% 
  pivot_longer(cols = c(China, Taiwan), #compressing our country columns
               names_to = "country", #creating new country column for the two countries
               values_to = "lifeExp") #directing values to new lifeExp column
```

# Part 2 - Multivariate Option 1: Life Expectancy and GDP per Capita over Time

## 2.1: Pivoting Data to the New Layout

```{r}
gapminder %>% 
  filter(country == "China" | country == "Taiwan") %>% 
  select(country, year, lifeExp, gdpPercap) %>% 
  pivot_wider(id_cols = year, #keeping year column
              names_from = country, #pulling values from country column for new columns
              names_sep = "--", #choosing separation the column components
              values_from = c(lifeExp, gdpPercap)) #pulling values from our measures for new columns
```

## 2.2: Pivoting Data Back to Original Format

```{r}
gapminder %>% 
  filter(country == "China" | country == "Taiwan") %>% 
  select(country, year, lifeExp, gdpPercap) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              names_sep = "--",
              values_from = c(lifeExp, gdpPercap)) %>% 
  pivot_longer(cols = -year, #selecting all columns except year to compress
               names_sep = "--", #identifying what separates the components embedded in the columns
               names_to = c(".value", "country")) #directing new columns and their content
```

Note to self: the order arguments are entered into names_to is important - reversing the order gives funky results. I'm guessing this is due to "lifeExp" and "gdpPercap" (which are the values I'm retaining into columns) being the first component in the column header, and the countries being the second component.

# Part 3 - Table Joins

```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```


## 3.1 - Adding Email Addresses to the Guestlist

First, lets exmamine our data files to help with visualizing the code.

```{r}
guest
email
```

```{r}
email_longer <- email %>% 
  separate_rows(guest, sep = ", ") #convert the email list from groups of guests to individual guests
guest %>% 
  left_join(email_longer, c("name" = "guest")) #join into the guest table using guest/name as common reference
```

I had some trouble getting the join to reconcile groups of parties in the email list and the lis of names in the guest list. To fix this, I first re-arranged the email list into individual names and joined the names column with the new, longer, guests column.

## 3.2 - Identifying Non-Guestlist Emails

```{r}
email_longer %>% 
  anti_join(guest, c("guest" = "name")) #removing shared rows between guest/name column from email table
```

## 3.3 - Making a Full Guestlist and Email List

```{r}
guest %>% 
  full_join(email_longer, c("name" = "guest")) #completely joining tables using name/guest columns as reference
```


