
---
title: "Gapminder_pt_2"
output: 
  html_document:
    keep_md: true 
    theme: cerulean
    toc: true
---

# STAT545 Assignment 2: Gapminder Further Exploration

```{r, warning=FALSE, echo=FALSE, message=FALSE}
library(gapminder)
library(tidyverse)
library(scales)
library(tsibble)
library(DT)
library(knitr)
library(kableExtra)
```

##Exercise 1: Basic dplyr

### 1.1 - Select Countries in 1970s

#### Canada, France, and New Zealand

```{r}
selectgapminder <- gapminder %>%
 filter(year >= 1970 & year <= 1979, 
   country == "Canada" | country == "France" | country == "New Zealand")
kable(selectgapminder)
```

### 1.2 - Select Countries and GDP Per Capita


```{r}
selectgapminder <- gapminder %>%
  filter(year >= 1970 & year <= 1979, 
  country == "Canada" | country == "France" | country == "New Zealand") %>%
  select (country, gdpPercap)
kable(selectgapminder)
```

### 1.3 - Changes in Life Expectancy

```{r}
selectgapminder <- gapminder %>%
  group_by(country)%>%
  arrange(year)%>%
  mutate(IncreaseInLifeExp = difference(lifeExp)) %>%
  arrange(IncreaseInLifeExp) %>%
  filter(IncreaseInLifeExp < 0)
datatable(selectgapminder)
```

### 1.4 - Each Country's Max GDP Per Capita

```{r}
selectdatatable <- gapminder %>% 
  group_by(country) %>%
  summarize(maxGdpPercap = max(gdpPercap))
datatable(selectgapminder)
```


### 1.5 - Canadian Life Expectancies compared to GDP Per Capita

```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(., aes(lifeExp, gdpPercap)) +
  geom_point(color = "dark blue") + 
  scale_y_log10("GDP per capita", labels = scales::dollar_format()) +
  xlab("Life Expectancy") +
  theme_minimal()
```

## Exercise 2: Explore Individual Variables with dplyr

```{r}
selectgapminder <- gapminder %>%
  group_by(country) %>%
  summarize(min_pop = min(pop), max_pop = max(pop), mean_pop = mean(pop), 
            median_pop= median(pop), sd_pop = sd(pop))
datatable(selectgapminder)
```

## Exercise 2: Explore Individual Variables with dplyr (cont.)

```{r}
gapminder %>%
  group_by(continent, country) %>%
  summarize(meanpop = mean(pop) / 1000000) %>%
  ggplot(., aes(continent, meanpop)) +
  geom_col(fill = "light blue") +
  ylab("Mean Population (millions)") +
  xlab("Continent") +
  theme_minimal()
```

## Exercise 2: Explore individual variables with dplyr (cont.)

```{r}
selectgapminder <- gapminder %>%
  group_by(continent, country) %>%
  summarize(n_countries = n_distinct(country)) %>%
  tally(n_countries)
kable(selectgapminder)
```


## Exercise 3: Exploring Various Plot Types

### Summary of iris data before jumping into plots

```{r}
library(datasets)
datatable(iris)
```

### Plot 1: Exploring petal width and petal length

#### This allows us to see, generally, which species are producing petals with the largest surface area

```{r}
iris %>%
  ggplot() +
  geom_point (aes(Petal.Width, Petal.Length, color=Species)) +
  xlab("Pedal Width") +
  ylab("Petal Length") +
  theme_minimal()
```

### Plot 2: Exploring Sepal Surface Area of the Three Species

#### Here we examine surface areas more directly between surface areas, this time exmaming sepal instead of petal. Of course, the surface area equations used in this example are for rectangles, and is therefore a crude approach that gives relative values.

```{r}
iris %>%
  ggplot() +aes(Species, iris$Sepal.Width*iris$Sepal.Length, fill = Species) + geom_boxplot() + ylab("Sepal Surface Area") + scale_fill_brewer(palette="RdBu") + theme_minimal()
```

## Recycling Bonus Exercise

## Recycling

#### The code is incorrect, instead of separately evaluating each country the code takes a little of both, resulting in the following 12 row table instead of 24:
```{r}
newgap <- filter(gapminder, country == c("Rwanda", "Afghanistan"))
datatable(newgap)
```
 
#### By separating the filter by 'or' statements, we get a table with all data points for all years for the 2 countries.

```{r}
newgap <- filter(gapminder, country == "Rwanda" | country == "Afghanistan")
datatable(newgap)
```