---
title: 'STAT545 Assignment 5: Factor and Figure Management'
output:
  html_document:
    toc: yes
    keep_md: yes
    theme: united
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, echo = FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(kableExtra))
```

# Exercise 1: The Value of 'here' Package - Superior File and Path Management

Broadly speaking, using the 'here' package is a more robust way of managing your and directories, paths, and files compared to conventional, manual management. Conventional set-ups like rm(list = ls()) doesn't truly clean your R session while setwd() lends itself to file paths designed exclusively for your computer and directory but not necessarily others. We can tackle some of these issues by setting up projects, but we still maximize functionality by using projects in tandem with the 'here' package. 'here' essentially does the dirty work of detecting root directories and using this information to navigate paths in a way that is neutral to your platform. Indeed, 'here' can operate seamlessly between operating systems like Mac and Windows, which have slightly different labelling systems for directory pathways. There is no need for another user to re-write code in order for the script to navigate their directory or their operating system. Further, if a project has sub-directories, 'here' is not restricted by a local working directory; it can use the project directory to seamlessly pull files from other locations or save files to different locations. If paths in the directory change, 'here' will still function correctly with no code changes required. 'here' will also allow files to run outside of their Rstudio project. Ultimately, a staple of data-driven science is reproducibility, and the 'here' package offers more seamless, reproducible workflow for ventures both within and between individuals.

# Exercise 2: Factor Management

Let's first take a look at our full gapminder data set for context in this exercise

```{r}
datatable(gapminder)
```

Now Let's Examine what categorical variables in our dataframe are factors

```{r}
class(gapminder$country)
class(gapminder$continent)
```

Therefore, both Country and Continent are factors. 

## Dropping Oceania

Before dropping Oceania, let's check how many oberervations we have so we can later verify their removal. This removal will happen in a new gapminder dataframe, so we'll make that now too and work with this new dataframe

```{r}
new_gapminder <- gapminder
levels(new_gapminder$continent)
nlevels(new_gapminder$continent)
```

Lit fam, we have 5 continents. Let's drop Oceania.

```{r}
keep_continents <- c("Americas", "Africa", "Asia", "Europe")

target_factors <- new_gapminder %>% 
  filter(continent %in% keep_continents)

dropped_gapminder <- target_factors %>% 
  droplevels()

levels(dropped_gapminder$continent)
nlevels(dropped_gapminder$continent)
```

## Re-ordering by Factors: Continental Min GDP

Let's organize our factors - in this case continent - by something other than the alphabet. How about minimum GDP per capita? First we'll find the min GDP, and then we'll do some sorting.

```{r}
gapminder %>% 
  select(continent, gdpPercap) %>%
  group_by(continent) %>%
  summarise(minGDP = min(gdpPercap)) %>%
  ggplot() +
  geom_col(aes(fct_reorder(continent, minGDP, max), minGDP, fill = continent)) +
  xlab("Continent") +
  ggtitle("Continental Minimum GDP per capita") +
  scale_y_continuous("GDP per capita", labels = scales::dollar_format()) +
  theme_minimal() +
  scale_fill_brewer() +
  theme(legend.position = "none")
```

# Exercise 3: File Input / Output

To reduce the size of the data we're working with, I chose to use Life Expectancy. I was born in 1997 in North America, so let's examine Life Expectancies in the Americas is the year 1997. First we'll filter the data set, then we'll write this into a 


```{r}
(gapminder_Amer <- gapminder %>% 
   filter(continent == "Americas", year == "1997") %>%
   select(country, lifeExp) %>%
   group_by(country)) %>%
  datatable()

write_csv(gapminder_Amer, here::here("Hw05", "lifeExpAmer.csv"))
(lifeExpAmer <- read_csv(here::here("Hw05", "lifeExpAmer.csv")))

lifeExpAmer %>%
  ggplot() +
  geom_col(aes(fct_reorder(country, lifeExp, max), lifeExp)) +
  scale_y_continuous("Life Expectancy") +
  coord_flip() +
  xlab("Country") +
  ggtitle("Life Expectancy of North and South American Countries in 1997 ") +
  theme_minimal()
```

