---
title: "STAT545 Assignment 3: Exploring Data with dplyr and ggplot pt. II"
output:
  html_document:
    toc: yes
    keep_md: yes
    theme: united
---

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, echo = FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(scales))
```

# Task X - blah: 

```{r}
gapminder
```


## X.1 - Tibble: 

### Defining 'Low Life Expectancy'

#### To establish thresholds for low and high life expectancy, lets summarise life expectancies of the global population
```{r}
gapminder %>% 
  summarise(mu = mean(lifeExp),
            sigma = sd(lifeExp)) 
```
#### Average life expactancy appears to be about 59.5 years. Low life expectancy can be defined as any life expectancy one standard deviation below the mean, representing the lower 32% of the population. As a absolute value, this is 46.5 years.

### Creating Tibble



## X.2 - Plot:

## X.3 - Discussion:

------

# Task X - blah: 

## X.1 - Tibble: 

## X.2 - Plot:

## X.3 - Discussion:

------

# Task X - blah: 

## X.1 - Tibble: 

## X.2 - Plot:

## X.3 - Discussion:

------

