---
title: "STAT545 Assignment 3: Exploring Data with dplyr and ggplot pt. II"
output:
  html_document:
    toc: yes
    keep_md: yes
    theme: united
---

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, echo = FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(scales))
```

# Task X - blah: 

```{r}
gapminder
```


## X.1 - Tibble: 

### Defining 'Low Life Expectancy'

#### To establish thresholds for low and high life expectancy, lets summarise life expectancies of the global population. LifeExp is related to year, therefore it may be misleading to ignore the year in obtaining these values to set our threshold. Using most recent year, for example, will yield the highest possible values. To compensate, we will examine life expactancy around the median year in the data set.

```{r}
median(gapminder$year)
```

```{r}
gapminder %>% 
  filter(year == "1982") %>% 
  summarise(mu = mean(lifeExp),
            sigma = sd(lifeExp)) 
```


#### Average life expactancy appears to be about 61.5 years in the year 1982. Low life expectancy can be defined as any life expectancy one standard deviation below the mean, representing the lower 32% of the population. As a absolute value, this is 50.7 years.

### Creating Tibble

```{r}
gapminder %>% 
  group_by(smallLifeExp = lifeExp < 50.7) %>% 
  arrange(desc(smallLifeExp))
```

```{r}
gapminder %>% 
  mutate(smallLifeExp = lifeExp < 50.7) %>% 
  select(year, country, continent, lifeExp, smallLifeExp) %>% 
  filter(smallLifeExp == 1) %>% 
  count(continent, year,)
```






















## X.2 - Plot:

## X.3 - Discussion:

------

# Task X - blah: 

## X.1 - Tibble: 

## X.2 - Plot:

## X.3 - Discussion:

------

# Task X - blah: 

## X.1 - Tibble: 

## X.2 - Plot:

## X.3 - Discussion:

------

