---
title: "Sample Presentation"
output: ioslides_presentation
--- 

```{r activations, include = FALSE}

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(DT))

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Gapminder Raw Data

```{r}
datatable(as_tibble(gapminder::gapminder), extensions = "FixedColumns")
```

## Life Expectency Across Time 

```{r}
ggplot(gapminder, aes(year, lifeExp)) + geom_point(alpha = 0.1, fill = "dark blue", colour = "dark blue") + theme_bw() + ylab("Life Expectancy")
```

## Let's Take a Look at Canada! | Life Expectancies {.build}

```{r}
CanLifeExp <- gapminder %>%
    filter(country == 'Canada') %>%
    group_by(year) %>%
    summarise(lifeExp = mean(lifeExp))
datatable(as_tibble(CanLifeExp))
```

## Let's Take a Look at Canada! (cont) | GDP {.build}

```{r}
CanGDP <- gapminder %>%
    filter(country == 'Canada') %>%
    group_by(year) %>%
    summarise(gdpPercap = mean(gdpPercap))
datatable(as_tibble(CanGDP))
```

## Let's Take a Look at Canada! (cont) | Canadian GDP over Time, Graphed {.build}


```{r}
ggplot(CanGDP, aes(year, gdpPercap)) + geom_line(linetype = "dashed", colour = "dark grey") + ylab("GDP Per Capita") + theme_minimal() + geom_point(colour = "red")
```

## Let's Take a Look at Canada! (cont) | Canadian Life Expectancy over Time, graphed {.build}


```{r}
ggplot(CanLifeExp, aes(year, lifeExp)) + geom_line(linetype = "dashed", colour = "dark grey") + ylab("Life Expectancy") + theme_minimal() + geom_point(colour = "red")
```

## How About the Americas? {.build}

```{r}
Am_lifeExp <- gapminder %>%
    filter(continent == 'Americas') %>%
    group_by(year)
datatable(as_tibble(Am_lifeExp))
```

## Americas' Life Expectancies Across Time {.build}

```{r}
ggplot(Am_lifeExp, aes(year, lifeExp, colour = country)) + geom_line() + ylab("Life Expectancy") + theme_minimal()
```


## Thanks for your time! :D
 
